---
title: "Package Datasets"
output: rmarkdown::html_vignette
number_sections: TRUE
link-citations: TRUE
bibliography: datasets.bib
csl: iso690-numeric-en.csl
vignette: >
  %\VignetteIndexEntry{Package Datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  webshot = "webshot",
  fig.width = 7.2,
  fig.height = 4.8
)
options(
  rmarkdown.html_vignette.check_title = FALSE
)

```

```{r setup, include=FALSE}

library(eprscope)
library(icons)
library(knitr)
library(htmlwidgets)
library(DT)

```

# Built-In Datasets

There are two important datasets which are built into the `{eprscope}` 📦. The first one is the nuclear isotope table summarizing the essential nuclei properties in order to analyze the EPR spectra with HF (hyperfine) structure. Whereas, the second one shows the solvent properties important for the variable temperature (VT) or double-resonance (ENDOR) EPR experiments.

## Nuclear Isotope Data Frame

This dataset was taken from the open source [EasySpin](https://easyspin.org/easyspin/documentation/isotopetable.html) package, reformatted and column of Larmor-frequencies in $\mathrm{MHz}$ at $0.35\,\mathrm{mT}$ were added for better orientation in double-resonance ENDOR spectra, see the details in `?isotopes_ds` . The data frame is not only used to analyze the ENDOR spectra however, it is essential for the simulations of EPR spectra and their fitting onto the experimental ones. See the documentation of the `eval_sim_EPR_iso()`, `eval_sim_EPR_isoFit()` as well as `eval_nu_ENDOR()`.

```{r nuclear-dataset}

# interactive data frame by `{DT}` package with option to select columns
# and to save table as `.csv`, `.pdf` or `.xlsx` format
DT::datatable(isotopes_ds,
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("colvis","csv","pdf","excel")
  )
)

```

## Solvent Properties

This is the key-feature when performing the EPR experiments.

```{r solvent-properties}

# similar interactive table like before
DT::datatable(solvents_ds,
  extensions = "Buttons",
  options = list(
    dom = "Bfrtip",
    buttons = c("colvis","csv","pdf","excel")
  )
)

```
