---
title: "Simulation of EPR Spectra"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js"
    toc: TRUE
number_sections: TRUE
link-citations: TRUE
bibliography: EPR_Simulations.bib
csl: iso690-numeric-en.csl
vignette: >
  %\VignetteIndexEntry{Simulation of EPR Spectra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7.2,
  fig.height = 4.8
)
options(
  rmarkdown.html_vignette.check_title = FALSE
)
```

$\require{mhchem}$

```{r setup, include = FALSE}
library(eprscope)
library(icons)
library(knitr)
library(patchwork)
library(ggplot2)
```

# Introduction

Quite often to figure out the HF structure from EPR spectra (see also `vignette("functionality")`) may be not so easy task even for a simple unpaired electron - interacting nuclei ($\ce{e^{-}-N}$) system as reported elsewhere [@Ramesova2013]. The reason is the interaction of the unpaired $\ce{e-}$ with all surrounding nuclei which spectral pattern may heavily overlap. Therefore, the theoretical predictions or computations of a spectrum (and its comparison with the experiment) represent an important step in the analysis of EPR spectra. However, such step requires an iterative process with modelling the above-described $\ce{e^{-}-N}$ system followed by a numerical simulations of the spectra to match the experimental ones. Commonly, quantum chemical calculations (usually DFT, see also `vignette("functionality")`) are involved in this process. As already mentioned in `README` and `vignette("introduction")` EPR simulations can be relatively easily performed by the great *EasySpin* toolbox within the proprietary MATLAB [@Stoll2007; @Easyspin2023]. Thus, also the `{eprscope}` 📦 contains functions to perform EPR simulations at least, in isotropic regime. It suggests that the molecules tumble/move extremely fast causing a total averaging of any anisotropic properties out. An EPR spectrum corresponding to the latter consists of a series of symmetric lines with equal widths. In such case it is characterized by the isotropic $g_{\mathrm{iso}}$-value, by the HF coupling/splitting constants $A_{\mathrm{iso}}/a_{\mathrm{iso}}$ as well as by the above-mentioned linewidth [@Stoll2007]. Many organic radicals exhibit such EPR spectra at room temperature and in solutions of low viscosities [@Gerson2003].

# Simplified Theoretical Description

The proper analytic expression for energies of magnetic dipole moments of one unpaired electron interacting with surrounding nuclei in the magnetic field (magnetic flux density, $B$) requires quantum mechanical operator formalism because the energies correspond to eigenvalues of the energy (so called Hamiltonian) operator [@WertzBoltonA; @WertzBoltonB; @WertzBoltonC]. However, for the sake of simplicity and illustration just consider the energy ($E$) relations. The essential formula how to calculate the energy of a magnetic dipole moment in magnetic field characterized by the magnetic flux density within the framework of "classic" magnetism corresponds to dot product like in Equation \@ref(eq:magnEnergy):

```{=tex}
\begin{equation}
E = - \vec{\mu}_{\mathrm{B}}^{}\,\vec{B}
(\#eq:magnEnergy)
\end{equation}
```
where the constant ... The total energy $E_{\mathrm{Tot}}$ for an isotropic system of one unpaired $\ce{e-}$ and the nuclei in its vicinity can be expressed by the following Equation \@ref(eq:eNenergy):

```{=tex}
\begin{equation}
E({\mathrm{Tot}}) = E({\mathrm{eZeeman}}) - \sum_k^n E_k({\mathrm{NZeeman}}) + \sum_k^n E_k({\mathrm{eNinteraction}})
(\#eq:eNenergy)
\end{equation}
```
# References {.unnumbered}
