---
title: "Simulation of EPR Spectra"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js"
    toc: TRUE
number_sections: TRUE
link-citations: TRUE
bibliography: EPR_Simulations.bib
csl: iso690-numeric-en.csl
vignette: >
  %\VignetteIndexEntry{Simulation of EPR Spectra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7.2,
  fig.height = 4.8
)
options(
  rmarkdown.html_vignette.check_title = FALSE
)
```

```{r setup, include = FALSE}
library(eprscope)
library(icons)
library(knitr)
library(patchwork)
library(ggplot2)
```

# Introduction

Quite often to figure out the HF structure from EPR spectra (see also `vignette("functionality")`) may be not so easy task even for a simple unpaired electron - interacting nuclei ($\ce{e^{-}-N}$) system as reported elsewhere [@Ramesova2013]. Therefore, the theoretical predictions or computations of a spectrum (and its comparison with the experiment) represent an important step in the analysis of EPR spectra. However, such step requires an iterative process with modelling the above-described $\ce{e^{-}-N}$ system followed by a numerical simulations of the spectra to match the experimental ones. Commonly, quantum chemical calculations (usually DFT, see also `vignette("functionality")`) are heavily involved in this process. As already mentioned in `README` and `vignette("introduction")` EPR simulations can be relatively easily performed by the are Blah Blah [@Easyspin2023]. Blah

# References {.unnumbered}
