---
title: "Basic Functionality Demonstration - Worflow to Process and Analyze EPR Spectrum of the Wuster's Blue"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js"
number_sections: TRUE
link-citations: TRUE
bibliography: functionality.bib
csl: iso690-numeric-en.csl
vignette: >
  %\VignetteIndexEntry{Basic Functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

$\require{mhchem}$

```{r setup, include = FALSE}
library(eprscope)
library(icons)
library(knitr)
```

# Introduction

BlahBlahBlah

and

-   Reading the universal ASCII data format (e.g. like `.asc`, `.txt`, `.csv`) and transforming into tables (data frames) incl. operations like intensity normalization and conversion between magnetic flux density (*B*) units (G ⟺ mT). Additionally, the functions enable the automatic reading of time series experiments like a kinetic one (e.g. also with the combination of *in situ* electrochemistry or photochemistry/irradiation) or reading and processing several EPR spectra data/files, at once, for the series of individual experiments/data files upon changing the quantities like time,microwave power, electrochemical potential, temperature, concentration...etc. Finally, reading the *MATLAB* (`.mat`) files and the instrumental parameter ones ( in case of *BRUKER* instruments, so far) and the corresponding parameter extraction can be performed as well.

-   Plotting the EPR spectra incl. time series/kinetics or their dependency on other quantities (like already mentioned above) as well as presentation of simulated spectra are provided by several functions based on either [ggplot2](https://ggplot2.tidyverse.org/) package/system (one of the most comprehensive system for data visualization) or interactive [plotly](https://plotly.com/r/) graphing library. Especially, the later may represent a valuable alternative to EPR instrumental software or other proprietary graphing systems, because it includes visualization tools like zooming, panning, data/values hovering and much more.

-   Quantifying the number of paramagnetic species (e.g. radicals) based on integrated forms of derivative EPR spectra and instrumental parameters. In this respect also baseline polynomial corrections are included to precisely evaluate the double integrals by numeric [trapezoidal method](https://mathworld.wolfram.com/TrapezoidalRule.html). In case of lower signal-to-noise ratios i.e. in noisy EPR spectra, they can be optimized by least-square fitting of simulated ones (e.g. from EasySpin) and single or double integrals can be evaluated using the optimized intensity of the simulated EPR spectrum. This is particularly advantageous in (time) series of many noisy EPR spectra. Such operations have been available only on EPR instruments/spectrometers, up to now.

-   Last but not least, there are functions to find extremes (maxima, minima), linewidths (incl. 𝚫*B*~pp~) as well as *g*-values directly from the spectral data. Eexperimentally (or by spectral simulations) determined EPR spectral parameters like *g*-values and hyperfine coupling (*A*)/splitting (*a*) constants are quite often compared to theoretically predicted by quantum chemical calculations (e.g. by [Density Functional Theory](https://gaussian.com/dft/)). Therefore
