---
title: MainTitle
author: 
  - name: 
    email: 
    orcid: 
    affiliations: 
      - name: 
        url:
date: today
---

::: content-hidden
$$
\require{mhchem}
$$
:::

```{r}
#| label: required-libraries
#| include: false

library(tidyverse)
library(eprscope)
library(patchwork)
library(knitr)
library(kableExtra)
library(gt)
library(epoxy)
library(tinytable)

```

```{r}
#| label: data-location-directories
#| include: false

raw.expr.data.loc <- file.path(".","Input_Data","EPR_RAW")
asc.expr.data.loc <- file.path(".","Input_Data","EPR_ASCII")
sim.easyspin.data.loc <- file.path(".","Input_Data","EasySpin_Simulations") 
fig.out.dir.loc <- file.path(".","_output","Figures")
tab.out.dir.loc <- file.path(".","_output","Tables")

```

# Experimental

This is experimental section ... electrolyte $0.2\,\mathrm{M}~\ce{TBAPF6}/\ce{CH3CN}$ ...

## Additional Chemistry

Acetonitrile $\ce{CH3CN}$, Dichloromethane $\ce{CH2Cl2}$ or corresponding radical $\ce{^.CH2Cl}$. Or Fenton reaction @eq-fenton

$$
\ce{H2O2 + Fe^2+ <=> Fe^3+ + ^.OH + ^{-}OH}
$$ {#eq-fenton}

{{< pagebreak >}}

# Computational Details

This section is dedicated to computations, processing and analysis of data ...

## Math & Physics

Schroedinger (@eq-schroedinger) 🙂 💻

$$
\begin{aligned}
\hat{H}\,|\psi\rangle & = \varepsilon\,|\psi\rangle  \\
\langle \psi|\,\hat{H}\,|\psi\rangle & = \varepsilon 
\end{aligned}
$$ {#eq-schroedinger}

{{< pagebreak >}}

# Results

Section summarizing all results + discussion ... Detailed description, see [@eprscope2024], ...

## Simulation of EPR Spectra

```{r}
#| label: fig-epr-simulation-example
#| echo: false
#| fig-cap: "This is example of the \\ce{TMPD^{.+}} simulated EPR spectrum."
#| warning: false
#| message: false

# simulation
sim.tmpd.spec <- eval_sim_EPR_iso(
  g.iso = 2.00303,
  instrum.params = c(Bcf = 3499.17,
                     Bsw = 120,
                     Npoints = 2401,
                     mwGHz = 9.814155),
  nuclear.system = list(
    list("14N", 2, 19.32),
    list("1H", 4, 5.50),
    list("1H", 12, 19.66)
  ),
  lineGL.DeltaB = list(0.44, 0.39),
  lineG.content = 0.4
)

# spectrum preview
sim.tmpd.spec$plot

```

@fig-epr-simulation-example illustrates the HF-structure of the EPR spectrum.

Figure out the $g$-value from the previous spectrum:

```{r}
#| label: epr-g-simulation-example
#| echo: true

eval_gFactor_Spec(sim.tmpd.spec$df,
  nu.GHz = 9.814155,
  B = "Bsim_G",
  Intensity = "dIeprSim_over_dB"
)

```

## Callouts 📓

::: callout-important
## Important Callout

Including something important
:::

{{< pagebreak >}}

# References {.unnumbered}
